---
title: "Applied Data Science with R Capstone project"
author: "Josue Afouda"
date: "2023-09-02"
output:
  pdf_document:
    toc: yes
  word_document:
    toc: yes
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(httr)
library(dplyr)
library(readr)
library(tidyr)
library(stringr)
library(fastDummies)
library(RSQLite)
library(ggplot2)
library(viridis)
library(tidymodels)
```

```{r, eval=FALSE}
# Libraries
library(rvest)
library(httr)
library(dplyr)
library(readr)
library(tidyr)
library(stringr)
library(fastDummies)
library(RSQLite)
library(ggplot2)
library(viridis)
library(tidymodels)
```


# Executive Summary

This document outlines a comprehensive data analysis and predictive modeling project focused on bike-sharing systems. The project is divided into several parts, each addressing key aspects of data collection, data wrangling, exploratory data analysis, predictive modeling, and the creation of a bike-sharing demand forecasting application.

Introduction

Bike-sharing systems have gained popularity as an environmentally friendly and convenient mode of transportation in urban areas. Understanding the factors influencing bike rental demand is crucial for optimizing bike-sharing operations. This document presents a step-by-step approach to collect, clean, analyze, and model data related to bike-sharing systems. The goal is to build predictive models that can forecast bike rental demand based on various factors, including weather conditions.

Part 1: Data Collection

In this section, data is collected from multiple sources, including web scraping a Global Bike-Sharing Systems Wiki Page and making API calls to OpenWeather for weather data. A custom function is created to gather 5-day weather forecasts for a list of cities. The collected data is saved as CSV files for further analysis.

Part 2: Data Wrangling

Data collected from various sources may require cleaning and standardization. This section focuses on data cleaning using regular expressions, transforming data into a suitable format, and preparing it for analysis. Special attention is given to standardizing column names and handling missing values.

Part 3: Exploratory Data Analysis (EDA)

EDA is performed using SQL queries and Tidyverse in R. Various tasks are executed, including determining data records, identifying patterns, summarizing statistics, and creating visualizations. The goal is to gain insights into the bike-sharing dataset and understand the relationships between variables.

Part 4: Predictive Modeling

This section involves building predictive models to forecast bike rental demand. Initially, a baseline linear regression model is created using weather-related variables. Subsequently, more complex models are developed, including polynomial regression models. Model performance is evaluated, and important features are identified to enhance prediction accuracy.

Part 5: Shiny Application Development

The final part focuses on creating an interactive bike-sharing demand forecasting application using R Shiny and Leaflet. This application allows users to input weather conditions and receive predictions for bike rental demand, enhancing the usability and accessibility of the predictive models.

Throughout this document, we provide detailed explanations and code examples to guide you through each step of the project, from data collection to application development. By the end of this project, you will have a comprehensive understanding of how to analyze and predict bike-sharing demand, which can be valuable for bike-sharing system operators and urban planners.



# Partie 1 : Data Collection

## Web scrape a Global Bike-Sharing Systems Wiki Page


```{r}
url <- "https://en.wikipedia.org/wiki/List_of_bicycle-sharing_systems"
# Get the root HTML node by calling the `read_html()` method with URL
root_html_node <- read_html(url)
```


```{r}
# Convert the bike-sharing system table into a dataframe
bike_sharing_systems <- html_table(root_html_node)[[1]]
head(bike_sharing_systems)
```

```{r}
# Summarize the dataframe
summary(bike_sharing_systems)
```

```{r}
# Export the dataframe into a csv file
write.csv(bike_sharing_systems, file = "raw_bike_sharing_systems.csv", row.names = FALSE)
```


## OpenWeather APIs Calls

### Get the current weather data for a city using OpenWeather API

The API base URL to get current weather is https://api.openweathermap.org/data/2.5/weather

```{r}
# URL for Current Weather API
current_weather_url <- 'https://api.openweathermap.org/data/2.5/weather'
```

Next, let's create a list to hold URL parameters for current weather API

```{r}
# need to be replaced by your real API key
your_api_key <- "7a4dc529d49a7fde9a79c07aa1eff515"
# Input `q` is the city name
# Input `appid` is your API KEY, 
# Input `units` are preferred units such as Metric or Imperial
current_query <- list(q = "Seoul", appid = your_api_key, units="metric")
```

Now we can make a HTTP request to the current weather API:

```{r}
response <- GET(current_weather_url, query = current_query)
response
```
To read the JSON HTTP response, you can use the content() function to parse it as a named list in R.

```{r}
json_result <- content(response, as = "parsed")
class(json_result)
json_result
```

We need to convert the named list to a data frame so that we can use data frame operations to process the data. Below is a simple example, which you may implement your own way to convert it to a data frame.


```{r}
# Create some empty vectors to hold data temporarily
weather <- c()
visibility <- c()
temp <- c()
temp_min <- c()
temp_max <- c()
pressure <- c()
humidity <- c()
wind_speed <- c()
wind_deg <- c()

# Now assign the values in the json_result list into different vectors
# $weather is also a list with one element, its $main element indicates the weather status such as clear or rain
weather <- c(weather, json_result$weather[[1]]$main)
# Get Visibility
visibility <- c(visibility, json_result$visibility)
# Get current temperature 
temp <- c(temp, json_result$main$temp)
# Get min temperature 
temp_min <- c(temp_min, json_result$main$temp_min)
# Get max temperature 
temp_max <- c(temp_max, json_result$main$temp_max)
# Get pressure
pressure <- c(pressure, json_result$main$pressure)
# Get humidity
humidity <- c(humidity, json_result$main$humidity)
# Get wind speed
wind_speed <- c(wind_speed, json_result$wind$speed)
# Get wind direction
wind_deg <- c(wind_deg, json_result$wind$deg)

# Combine all vectors as columns of a data frame
weather_data_frame <- data.frame(weather=weather, 
                                 visibility=visibility, 
                                 temp=temp, 
                                 temp_min=temp_min, 
                                 temp_max=temp_max, 
                                 pressure=pressure, 
                                 humidity=humidity, 
                                 wind_speed=wind_speed, 
                                 wind_deg=wind_deg)

# Check the generated data frame
print(weather_data_frame)
```
Généralisons le processus ci-dessous en écrivant une fonction qui retourne une dataframe contenant les informations sur la météo actuelle de plusieurs villes :

```{r}
# Fonction pour récupérer automatiquement des informations sur la météo actuelle de plusieurs villes
get_weather_info <- function(cities, api_key, current_weather_url = 'https://api.openweathermap.org/data/2.5/weather') {
  weather_data <- data.frame()

  for (city in cities) {
    # Construct the API query for each city
    current_query <- list(q = city, appid = api_key, units = "metric")
    
    # Make a HTTP request to the current weather API
    response <- GET(current_weather_url, query = current_query)
    
    # Check if the request was successful
    if (http_type(response) == "application/json") {
      # Read the JSON HTTP response
      json_result <- content(response, as = "parsed")
      
      # Extract relevant weather information
      city_data <- data.frame(
        City = city,
        Weather_Description = json_result$weather[[1]]$description,
        Visibility = json_result$visibility,
        Temperature = json_result$main$temp,
        Temperature_Min = json_result$main$temp_min,
        Temperature_Max = json_result$main$temp_max,
        Pressure = json_result$main$pressure,
        Humidity = json_result$main$humidity,
        Wind_Speed = json_result$wind$speed,
        wind_Deg = json_result$wind$deg
      )
      
      # Append the data to the weather_data dataframe
      weather_data <- bind_rows(weather_data, city_data)
    } else {
      cat(paste("Error fetching data for", city, "\n"))
    }
  }
  
  return(weather_data)
}
```


Appliquons cette fonction :

```{r}
# Example usage:
your_api_key <- "7a4dc529d49a7fde9a79c07aa1eff515"
cities <- c("Seoul", "New York", "Paris")
weather_df <- get_weather_info(cities, your_api_key)
weather_df
```



### Get 5-day weather forecasts for a list of cities using the OpenWeather API

Now you should be familiar with the usage of OpenWeather API. Next, you need to complete a task to get 5-day weather forecasts for a list of cities

TODO: Write a function to return a data frame containing 5-day weather forecasts for a list of cities.


```{r}
get_weather_forecaset_by_cities <- function(cities, api_key) {
  
  forecast_data <- data.frame()

  for (city in cities) {
    # Construct the API query for each city
    forecast_query <- list(q = city, appid = api_key, units = "metric")
    
    # Make a HTTP request to the 5-day forecast API
    response <- GET('https://api.openweathermap.org/data/2.5/forecast', query = forecast_query)
    
    # Check if the request was successful
    if (http_type(response) == "application/json") {
      # Read the JSON HTTP response
      json_result <- content(response, as = "parsed")
      
      # Extract relevant forecast information (first 5 entries, one per day)
      forecast_entries <- json_result$list[1:5]
      
      for (entry in forecast_entries) {
        date <- entry$dt_txt
        weather_description <- entry$weather[[1]]$description
        visibility <- entry$visibility
        temperature <- entry$main$temp
        pressure <- entry$main$pressure
        humidity <- entry$main$humidity
        wind_speed <- entry$wind$speed
        wind_deg <- entry$wind$deg
        
        # Create a data frame for each forecast entry
        forecast_data <- bind_rows(forecast_data, data.frame(
          City = city,
          Date = date,
          Weather_Description = weather_description,
          Visibility = visibility,
          Temperature = temperature,
          Pressure = pressure,
          Humidity = humidity,
          Wind_Speed = wind_speed,
          Wind_Deg = wind_deg
        ))
      }
    } else {
      cat(paste("Error fetching data for", city, "\n"))
    }
  }
  
  return(forecast_data)
}
```


Appliquons la fonction :

```{r}
# Example usage:
your_api_key <- "7a4dc529d49a7fde9a79c07aa1eff515"
cities <- cities <- c("Seoul", "Washington, D.C.", "Paris", "Suzhou")
cities_weather_df <- get_weather_forecaset_by_cities(cities, your_api_key)
cities_weather_df
```

```{r}
# Write cities_weather_df to `raw_cities_weather_forecast.csv`
write.csv(cities_weather_df, "raw_cities_weather_forecast.csv", row.names=FALSE)
```


## Download datasets as csv files from cloud storage

```{r}
# Download several datasets

# Download some general city information such as name and locations
url <- "https://cf-courses-data.s3.us.cloud-object-storage.appdomain.cloud/IBMDeveloperSkillsNetwork-RP0321EN-SkillsNetwork/labs/datasets/raw_worldcities.csv"
# download the file
download.file(url, destfile = "raw_worldcities.csv")

# Download a specific hourly Seoul bike sharing demand dataset
url <- "https://cf-courses-data.s3.us.cloud-object-storage.appdomain.cloud/IBMDeveloperSkillsNetwork-RP0321EN-SkillsNetwork/labs/datasets/raw_seoul_bike_sharing.csv"
# download the file
download.file(url, destfile = "raw_seoul_bike_sharing.csv")
```

# Partie 2 : Data Wrangling

Dans la partie 1 sur la collecte de données, vous avez collecté des ensembles de données brutes provenant de plusieurs sources différentes. Dans cette deuxième partie, vous devez effectuer des tâches de nettoyage des données afin d'améliorer leur qualité.

## Nettoyage des données à l'aide d'expressions régulièress

Vous utiliserez des expressions régulières, ainsi que le package stringr (qui fait partie de Tidyverse), pour nettoyer les données des systèmes de partage de vélos que vous avez précédemment récupérées sur la page wiki :

```{r}
head(bike_sharing_systems)
```

Pour résumer, vous utiliserez `stringr` (qui fait partie de `tidyverse`) et des expressions régulières pour effectuer les tâches de nettoyage des données suivantes :

   - Standardiser les noms de colonnes pour tous les ensembles de données collectés ;
   
   - Supprimer les liens de référence indésirables de l'ensemble de données des systèmes de partage de vélos récupéré ;
   
   - Extraire uniquement la valeur numérique des annotations de texte indésirables.

### Standardiser les noms de colonnes pour tous les ensembles de données collectés

Voici les 4 ensembles de données qui ont été collectés dans la partie 1 :

- `raw_bike_sharing_systems.csv`:  Une liste des systèmes de location de vélos actifs à travers le monde ;

- `raw_cities_weather_forecast.csv`: Prévisions météo sur 5 jours pour une liste de villes, à partir de l'API OpenWeather ;

- `raw_worldcities.csv`: Une liste d'informations sur les grandes villes (telles que le nom, la latitude et la longitude) à travers le monde ;

- `raw_seoul_bike_sharing.csv`: Informations météorologiques (température, humidité, vitesse du vent, visibilité, point de rosée, rayonnement solaire, chutes de neige, précipitations), nombre de vélos loués par heure et informations sur la date, provenant des systèmes de location de vélos de Séoul.

Pour améliorer la lisibilité des ensembles de données par les systèmes humains et informatiques, nous devons d'abord standardiser les noms de colonnes des ensembles de données ci-dessus en utilisant la convention de dénomination suivante : 

* *Les noms de colonnes doivent être en MAJUSCULES* ; 

* *Le séparateur de mots doit être un trait de soulignement, comme dans `COLUMN_NAME`*.

```{r}
dataset_list <- c(
  'raw_bike_sharing_systems.csv', 
  'raw_seoul_bike_sharing.csv', 
  'raw_cities_weather_forecast.csv', 
  'raw_worldcities.csv'
)

for (dataset_name in dataset_list){
    # Read dataset
    dataset <- read_csv(dataset_name)
    
    # Standardized its columns:
    # Convert all column names to uppercase
    names(dataset) <- toupper(names(dataset))
    
    # Replace any white space separators by underscores, using the str_replace_all function
    names(dataset) <- gsub(" ", "_", names(dataset))
    
    # Save the dataset 
    write.csv(dataset, dataset_name, row.names=FALSE)
}
```


Let's Read the resulting datasets back and check whether their column names follow the naming convention:

```{r, warning=FALSE, message=FALSE}
for (dataset_name in dataset_list){
    # Print a summary for each data set to check whether the column names were correctly converted
  print(dataset_name)
  print(summary(read_csv(dataset_name)))
  print("----------------------------------------")
}
```

### Process the web-scraped bike sharing system dataset

By now we have standardized all column names. Next, we will focus on cleaning up the values in the web-scraped bike sharing systems dataset.

```{r}
# First load the dataset
bike_sharing_df <- read_csv("raw_bike_sharing_systems.csv")

# Print its head
head(bike_sharing_df)
```

Even from the first few rows, you can see there is plenty of undesireable embedded textual content, such as the reference link included in `Melbourne[12]`.

In this project, let's only focus on processing the following revelant columns (feel free to process the other columns for more practice):

 - `COUNTRY`: Country name 
 
 - `CITY`: City name
 
 - `SYSTEM`: Bike-sharing system name
 
 - `BICYCLES`: Total number of bikes in the system
 
 
```{r}
# Select the four columns
sub_bike_sharing_df <- bike_sharing_df %>% 
  select(COUNTRY, CITY, SYSTEM, BICYCLES)

# Let's see the types of the selected column
sub_bike_sharing_df %>% 
    summarize_all(class) %>%
    gather(variable, class)
```

They are all interpreted as character columns, but we expect the `BICYCLES` column to be of numeric type. Let's see why it wasn't loaded as a numeric column - possibly some entries contain characters. Let's create a simple function called `find_character` to check that.

```{r}
# grepl searches a string for non-digital characters, and returns TRUE or FALSE
# if it finds any non-digital characters, then the bicyle column is not purely numeric
find_character <- function(strings) grepl("[^0-9]", strings)
```

Let's try to find any elements in the `Bicycles` column containing non-numeric characters:

```{r}
sub_bike_sharing_df %>% 
    select(BICYCLES) %>% 
    filter(find_character(BICYCLES)) 
```

As you can see, many rows have non-numeric characters, such as `32 (including 6 rollers) [162]` and `1000[253]`. This is actually very common for a table scraped from Wiki when no input validation is enforced. 

Later, you will use regular expressions to clean them up.

Next, let's take a look at the other columns, namely `COUNTRY`, `CITY`, and `SYSTEM`, to see if they contain any undesired reference links, such as in `Melbourne[12]`.

```{r}
# Define a 'reference link' character class, 
# `[A-z0-9]` means at least one character 
# `\\[` and `\\]` means the character is wrapped by [], such as for [12] or [abc]
ref_pattern <- "\\[[A-z0-9]+\\]"
find_reference_pattern <- function(strings) grepl(ref_pattern, strings)
```


```{r}
# Check whether the COUNTRY column has any reference links
sub_bike_sharing_df %>% 
    select(COUNTRY) %>% 
    filter(find_reference_pattern(COUNTRY))
```

Ok, looks like the COUNTRY column is clean. Let's check the CITY column.

```{r}
# Check whether the CITY column has any reference links
sub_bike_sharing_df %>% 
    select(CITY) %>% 
    filter(find_reference_pattern(CITY))
```

Hmm, looks like the CITY column has some reference links to be removed. Next, let's check the SYSTEM column.

```{r}
# Check whether the System column has any reference links
sub_bike_sharing_df %>% 
    select(SYSTEM) %>% 
    filter(find_reference_pattern(SYSTEM)) 
```

So the `SYSTEM` column also has some reference links.

After some preliminary investigations, we identified that the `CITY` and `SYSTEM` columns have some undesired reference links, and the `BICYCLES` column has both reference links and some 
textual annotations.

Next, you need to use regular expressions to clean up the unexpected reference links and text annotations in numeric values.

### Remove undesired reference links using regular expressions

Let's Write a custom function using `stringr::str_replace_all` to replace all reference links with an empty character for columns `CITY` and `SYSTEM`:

```{r}
# Define a custom function to remove reference links
remove_ref <- function(strings) {
  # Define a pattern to match reference links, e.g., [1]
  ref_pattern <- "\\[[0-9]+\\]"
  
  # Replace all matched substrings with an empty character using str_replace_all()
  cleaned_strings <- str_replace_all(strings, ref_pattern, "")
  
  # Optionally, you can trim the resulting strings to remove any leading or trailing whitespace
  cleaned_strings <- str_trim(cleaned_strings) # remove whitespace
  
  # Return the cleaned character vector
  return(cleaned_strings)
}
```

Appliquons la fonctions `remove_ref()` sur les colonnes `CITY` et `SYSTEM` :

```{r}
result <- sub_bike_sharing_df %>% 
  mutate(
    CITY = remove_ref(CITY),
    SYSTEM = remove_ref(SYSTEM),
    
  )

# check whether all reference links are removed
result %>% 
    select(CITY, SYSTEM) %>% 
    filter(find_reference_pattern(CITY) | find_reference_pattern(SYSTEM))
```


### Extract the numeric value using regular expressions

Let's Write a custom function using `stringr::str_extract` to extract the first digital substring match and convert it into numeric type For example, extract the value '32' from `32 (including 6 rollers) [162]`.

```{r}
# Extract the first number
extract_num <- function(columns) {
  # Define a pattern to match the first numeric substring
  digitals_pattern <- "\\d+"  # Matches one or more digits
  
  # Find the first match using str_extract
  first_numeric <- str_extract(columns, digitals_pattern)
  
  # Convert the result to numeric using the as.numeric() function
  numeric_value <- as.numeric(first_numeric)
  
  # Return the numeric value
  return(numeric_value)
}
```


Appliquons la fonction `extract_num()` à la colonne ` :

```{r}
result <- result %>%
  mutate(BICYCLES = extract_num(BICYCLES))

summary(result$BICYCLES)
```

Sauvegardons, dans un fichier CSV, l'ensemble de données nettoyées des systèmes de location de vélos :

```{r}
write.csv(result, file = "bike_sharing_systems.csv", row.names = FALSE)
```


## Nettoyage des données avec dplyr

In this lab, you will focus on wrangling the Seoul bike-sharing demand historical dataset. This is the core dataset to build a predictive model later.

It contains the following columns:

- `DATE` : Year-month-day

- `RENTED BIKE COUNT`- Count of bikes rented at each hour

- `HOUR`- Hour of he day

- `TEMPERATURE` - Temperature in Celsius

- `HUMIDITY` - Unit is `%`

- `WINDSPEED` - Unit is `m/s`

- `VISIBILITY` - Multiplied by 10m

- `DEW POINT TEMERATURE` - The temperature to which the air would have to cool down in order to reach saturation, unit is Celsius

- `SOLAR RADIATION` - MJ/m2

- `RAINFALL` - mm

- `SNOWFALL` - cm

- `SEASONS` - Winter, Spring, Summer, Autumn

- `HOLIDAY` - Holiday/No holiday

- `FUNCTIONAL DAY` - NoFunc(Non Functional Hours), Fun(Functional hours)


For this dataset, you will be asked to use `tidyverse` to perform the following data wrangling tasks:

- **TASK: Detect and handle missing values**

- **TASK: Create indicator (dummy) variables for categorical variables**

- **TASK: Normalize data**

Let's start!


```{r, warning=FALSE, message=FALSE}
seoul_bike_sharing_df <- read_csv("raw_seoul_bike_sharing.csv")
dim(seoul_bike_sharing_df)
summary(seoul_bike_sharing_df)
```


From the summary, we can observe that:

Columns `RENTED_BIKE_COUNT`, `TEMPERATURE`, `HUMIDITY`, `WIND_SPEED`, `VISIBILITY`, `DEW_POINT_TEMPERATURE`, `SOLAR_RADIATION`, `RAINFALL`, `SNOWFALL` are numerical variables/columns and require normalization. Moreover, `RENTED_BIKE_COUNT` and `TEMPERATURE` have some missing values (NA's) that need to be handled properly.

`SEASONS`, `HOLIDAY`, `FUNCTIONING_DAY` are categorical variables which need to be converted into indicator columns or dummy variables.
Also, `HOUR` is read as a numerical variable but it is in fact a categorical variable with levels ranging from 0 to 23.

Now that you have some basic ideas about how to process this bike-sharing demand dataset, let's start working on it!

### Detect and handle missing values

 The `RENTED_BIKE_COUNT` column has about 295 missing values, and `TEMPERATURE` has about 11 missing values. Those missing values could be caused by not being recorded, or from malfunctioning bike-sharing systems or weather sensor networks. In any cases, the identified missing values have to be properly handled.
 
 Considering `RENTED_BIKE_COUNT` is the response variable/dependent variable, i.e., we want to predict the `RENTED_BIKE_COUNT` using other predictor/independent variables later, and we normally can not allow missing values for the response variable, so missing values for response variable must be either dropped or imputed properly. 

We can see that `RENTED_BIKE_COUNT` only has about 3% missing values (295 / 8760). As such, you can safely drop any rows whose `RENTED_BIKE_COUNT` has missing values.

```{r}
seoul_bike_sharing_df <- seoul_bike_sharing_df %>%
  drop_na(RENTED_BIKE_COUNT)

dim(seoul_bike_sharing_df)
```
Now that you have handled  missing values in the `RENTED_BIKE_COUNT` variable, let's continue processing missing values for the `TEMPERATURE` column.

Unlike the `RENTED_BIKE_COUNT` variable, `TEMPERATURE` is not a response variable. However, it is still an important predictor variable - as you could imagine, there may be a positve correlation between `TEMPERATURE` and `RENTED_BIKE_COUNT`. For example, in winter time with lower temperatures, people may not want to ride a bike, while in summer with nicer weather, they are more likely to rent a bike.

How do we handle missing values for `TEMPERATURE`? We could simply remove the rows but it's better to impute them because `TEMPERATURE` should be relatively easy and reliable to estimate statistically.

Let's first take a look at the missing values in the TEMPERATURE column.

```{r}
seoul_bike_sharing_df %>%
  filter(is.na(TEMPERATURE))
```

It seems that all of the missing values for `TEMPERATURE` are found in `SEASONS == Summer`, so it is reasonable to impute those missing values with the summer average temperature.

Let's Impute missing values for the TEMPERATURE column using its mean value for the Summer season.

```{r}
# Calculate the summer average temperature
avg_summer <- seoul_bike_sharing_df %>%
  filter(SEASONS == "Summer") %>%
  summarise(avg = mean(TEMPERATURE, na.rm = TRUE)) %>%
  pull(avg)

avg_summer
```

```{r}
# Impute missing values for TEMPERATURE column with summer average temperature

seoul_bike_sharing_df <- seoul_bike_sharing_df %>%
  mutate(
    TEMPERATURE = ifelse(is.na(TEMPERATURE), avg_summer, TEMPERATURE)
  )

# Print the summary of the dataset again to make sure no missing values in all columns
summary(seoul_bike_sharing_df)
```

```{r}
# Save the dataset as `seoul_bike_sharing.csv`
write.csv(
  seoul_bike_sharing_df,
  file = "seoul_bike_sharing.csv",
  row.names = FALSE
)
```

### Create indicator (dummy) variables for categorical variables

Regression models can not process categorical variables directly, thus we need to convert them into indicator variables.

In the seoul bike-sharing demand dataset, `SEASONS`, `HOLIDAY`, `FUNCTIONING_DAY` are categorical variables.Also, `HOUR` is read as a numerical variable but it is in fact a categorical variable with levels ranged from 0 to 23.

```{r}
# Convert HOUR column from numeric into character first:
seoul_bike_sharing_df <- seoul_bike_sharing_df %>%
  mutate(HOUR = as.character(HOUR))
```


`SEASONS`, `HOLIDAY`, `FUNCTIONING_DAY`,  `HOUR` are all character columns now and are ready to be converted into indicator variables.

For example, `SEASONS` has four categorical values: `Spring`, `Summer`, `Autumn`, `Winter`. We thus need to create four indicator/dummy variables `Spring`, `Summer`, `Autumn`, and `Winter` which only have the value 0 or 1.

So, given a data entry with the value `Spring` in the `SEASONS` column, the values for the four new columns `Spring`, `Summer`, `Autumn`, and `Winter` will be set to 1 for `Spring` and 0 for the others:

|Spring|Summer|Autumn|Winter|
|----- |------|------|------|
|     1|     0|     0|     0| 


```{r}
# Convert SEASONS, HOLIDAY, FUNCTIONING_DAY, and HOUR columns into dummy variables

seoul_bike_sharing_df <- seoul_bike_sharing_df %>%
  dummy_cols(
    select_columns = c("SEASONS", "HOLIDAY", "FUNCTIONING_DAY", "HOUR"),
    remove_selected_columns = TRUE
  )

# Print the dataset summary again to make sure the indicator columns are created properly
summary(seoul_bike_sharing_df)
```

```{r}
# Save the dataset as `seoul_bike_sharing_converted.csv`
write.csv(
  seoul_bike_sharing_df, 
  file = "seoul_bike_sharing_converted.csv",
  row.names = FALSE
)
```


### Normalize data

Let's perform normalization on numeric columns to transfer them into a similar range using Min-Max normalization.

```{r}
# Define a function for min-max normalization
min_max_normalize <- function(x) {
  (x - min(x)) / (max(x) - min(x))
}

# Apply min-max normalization to numeric columns using dplyr
seoul_bike_sharing_df <- seoul_bike_sharing_df %>%
  mutate(across(where(is.numeric), min_max_normalize))

summary(seoul_bike_sharing_df)
```

```{r}
# Save the dataset as `seoul_bike_sharing_converted_normalized.csv`
write.csv(
  seoul_bike_sharing_df, 
  file = "seoul_bike_sharing_converted_normalized.csv",
  row.names = FALSE
)
```


# Create tables and load data in a SQLite database

## Create tables

Commençons d'abord par créer une database et établissons la connexion :

```{r}
# Establish database connection
  # Ce code crée une database SQLite nommée "Capstone_Project.sqlite"
  # et établit la connexion à cette database.
conn <- dbConnect(SQLite(),"Capstone_Project.sqlite")
conn
```
Créons les tables :

```{r}
# WORLD_CITIES
  # ce code crée la table WORLD_CITIES
df1 <- dbExecute(conn, 
                 "CREATE TABLE WORLD_CITIES (
                                          CITY	VARCHAR(50),
                                          CITY_ASCII VARCHAR(50),
                                          LAT DECIMAL(20,2),
                                          LNG DECIMAL(20,2),
                                          COUNTRY	VARCHAR(50),
                                          ISO2 VARCHAR(5),
                                          ISO3 VARCHAR(5),
                                          ADMIN_NAME VARCHAR(100),	
                                          CAPITAL	VARCHAR(50),
                                          POPULATION BIGINT,
                                          ID BIGINT NOT NULL
                                      )", 
                 errors=FALSE
)

if (df1 == -1){
  cat ("An error has occurred.\n")
  msg <- odbcGetErrMsg(conn)
  print (msg)
} else {
  cat ("Table was created successfully.\n")
}

# BIKE_SHARING_SYSTEMS
  # ce code crée la table BIKE_SHARING_SYSTEMS
df2 <- dbExecute(conn, 
                 "CREATE TABLE BIKE_SHARING_SYSTEMS (
                                                COUNTRY	VARCHAR(20),
                                                CITY VARCHAR(87),
                                                SYSTEM VARCHAR(40),
                                                BICYCLES VARCHAR(5)
                                            )", 
                 errors=FALSE
)

if (df2 == -1){
  cat ("An error has occurred.\n")
  msg <- odbcGetErrMsg(conn)
  print (msg)
} else {
  cat ("Table was created successfully.\n")
}

# CITIES_WEATHER_FORECAST
  # ce code crée la table CITIES_WEATHER_FORECAST
df3 <- dbExecute(conn, 
                 "CREATE TABLE CITIES_WEATHER_FORECAST (
                                            CITY VARCHAR(16),
                                            WEATHER VARCHAR(6),
                                            VISIBILITY SMALLINT,
                                            TEMP DECIMAL(6,2),
                                            TEMP_MIN DECIMAL(6,2),
                                            TEMP_MAX DECIMAL(6,2),
                                            PRESSURE SMALLINT,
                                            HUMIDITY SMALLINT,
                                            WIND_SPEED DECIMAL(6,2),
                                            WIND_DEG SMALLINT,
                                            SEASON VARCHAR(6),
                                            FORECAST_DATETIME TIMESTAMP
                                        )", 
                 errors=FALSE
)

if (df3 == -1){
  cat ("An error has occurred.\n")
  msg <- odbcGetErrMsg(conn)
  print (msg)
} else {
  cat ("Table was created successfully.\n")
}


# SEOUL_BIKE_SHARING
  # ce code crée la table SEOUL_BIKE_SHARING
df4 <- dbExecute(conn, 
                 "CREATE TABLE SEOUL_BIKE_SHARING (
                                        DATE VARCHAR(30),
                                        RENTED_BIKE_COUNT SMALLINT,
                                        HOUR SMALLINT,
                                        TEMPERATURE DECIMAL(4,1),
                                        HUMIDITY SMALLINT,
                                        WIND_SPEED DECIMAL(3,1),
                                        VISIBILITY SMALLINT,
                                        DEW_POINT_TEMPERATURE DECIMAL(4,1),
                                        SOLAR_RADIATION DECIMAL(5,2),
                                        RAINFALL DECIMAL(3,1),
                                        SNOWFALL DECIMAL(3,1),
                                        SEASONS	VARCHAR(10),
                                        HOLIDAY	VARCHAR(20),
                                        FUNCTIONING_DAY VARCHAR(5)
                                    )", 
                 errors=FALSE
)

if (df4 == -1){
  cat ("An error has occurred.\n")
  msg <- odbcGetErrMsg(conn)
  print (msg)
} else {
  cat ("Table was created successfully.\n")
}
```

```{r}
# Liste des tables présentes dans la database
dbListTables(conn)
```

## Load the data in the database

```{r}
# Lire les fichiers de données en tant que dataframes
world_cities_cleaned <- read.csv("world_cities_cleaned.csv")
head(world_cities_cleaned, 3)

bike_sharing_systems_cleaned <- read.csv("bike_sharing_systems_cleaned.csv")
head(bike_sharing_systems_cleaned, 3)

cities_weather_forecast_cleaned <- read.csv(
  "cities_weather_forecast_cleaned.csv",
  colClasses = c(FORECAST_DATETIME="character")
)
head(cities_weather_forecast_cleaned, 3)

seoul_bike_sharing_cleaned <- read.csv("seoul_bike_sharing_cleaned.csv")
head(seoul_bike_sharing_cleaned, 3)
```

```{r}
# Chargement des données dans les tables
dbWriteTable(
  conn, 
  "WORLD_CITIES", 
  world_cities_cleaned, 
  overwrite = TRUE, 
  header = TRUE
)

dbWriteTable(
  conn, 
  "BIKE_SHARING_SYSTEMS", 
  bike_sharing_systems_cleaned, 
  overwrite=TRUE, 
  header = TRUE
)

dbWriteTable(
  conn, 
  "CITIES_WEATHER_FORECAST", 
  cities_weather_forecast_cleaned, 
  overwrite=TRUE, 
  header = TRUE
)

dbWriteTable(
  conn, 
  "SEOUL_BIKE_SHARING", 
  seoul_bike_sharing_cleaned, 
  overwrite=TRUE, 
  header = TRUE
)
```


# Partie 3 : Analyse exploratoire des donnés

## Analyse exploratoire des donnés avec SQL

### Task 1: Determine how many records are in the seoul_bike_sharing dataset

```{r}
dbGetQuery(conn, "SELECT count(*) FROM SEOUL_BIKE_SHARING")
```


### Task 2: Determine how many hours had non-zero rented bike count

```{r}
dbGetQuery(
  conn,
  "SELECT COUNT(*) AS NonZeroRentedHours
  FROM SEOUL_BIKE_SHARING
  WHERE RENTED_BIKE_COUNT > 0;"
)
```


### Task 3: Query the weather forecast for Seoul over the next 3 hours

Recall that the records in the CITIES_WEATHER_FORECAST dataset are 3 hours apart, so we just need the first record from the query.

```{r}
dbGetQuery(
  conn,
  "
  SELECT *
  FROM CITIES_WEATHER_FORECAST
  WHERE FORECAST_DATETIME BETWEEN (
    SELECT MIN(FORECAST_DATETIME) FROM CITIES_WEATHER_FORECAST
  ) AND (
    SELECT DATETIME(MIN(FORECAST_DATETIME), '+3 HOURS') FROM CITIES_WEATHER_FORECAST
  );

  "
)
```


### Task 4: Find which seasons are included in the seoul bike sharing dataset

```{r}
dbGetQuery(
  conn,
  "
  SELECT DISTINCT(SEASONS)
  FROM SEOUL_BIKE_SHARING
  "
)
```

### Task 5: Find the first and last dates in the Seoul Bike Sharing dataset

```{r}
dbGetQuery(
  conn,
  "
  SELECT MIN(DATE), MAX(DATE)
  FROM SEOUL_BIKE_SHARING
  "
)
```


### Task 6: determine which date and hour had the most bike rentals

```{r}
dbGetQuery(
  conn,
  "
  SELECT DATE, HOUR, MAX(RENTED_BIKE_COUNT) AS MAX_RENTALS
  FROM SEOUL_BIKE_SHARING
  GROUP BY DATE, HOUR
  ORDER BY MAX_RENTALS DESC
  LIMIT 1
  "
)
```


### Task 7: Determine the average hourly temperature and the average number of bike rentals per hour over each season

List the top ten results by average bike count.


```{r}
dbGetQuery(
  conn,
  "
  SELECT
      SEASONS,
      HOUR,
      AVG(TEMPERATURE) AS AVG_TEMPERATURE,
      AVG(RENTED_BIKE_COUNT) AS AVG_BIKE_COUNT
  FROM SEOUL_BIKE_SHARING
  GROUP BY SEASONS, HOUR
  ORDER BY AVG_BIKE_COUNT DESC
  LIMIT 10
  "
)
```

### Task 8: Find the average hourly bike count during each season

Also include the minimum, maximum, and standard deviation of the hourly bike count for each season.


```{r}
dbGetQuery(
  conn,
  "
  SELECT
    SEASONS,
    HOUR,
    AVG(RENTED_BIKE_COUNT) AS AVG_BIKE_COUNT,
    MIN(RENTED_BIKE_COUNT) AS MIN_BIKE_COUNT,
    MAX(RENTED_BIKE_COUNT) AS MAX_BIKE_COUNT
  FROM SEOUL_BIKE_SHARING
  GROUP BY SEASONS, HOUR
  ORDER BY SEASONS, HOUR
  "
)
```

### Task 9: On average, what were the TEMPERATURE, HUMIDITY, WIND_SPEED, VISIBILITY, DEW_POINT_TEMPERATURE, SOLAR_RADIATION, RAINFALL, and SNOWFALL per season?

Include the average bike count as well , and rank the results by average bike count so you can see if it is correlated with the weather at all.


```{r}
dbGetQuery(
  conn,
  "
  SELECT
    SEASONS,
    AVG(TEMPERATURE) AS AVG_TEMPERATURE,
    AVG(HUMIDITY) AS AVG_HUMIDITY,
    AVG(WIND_SPEED) AS AVG_WIND_SPEED,
    AVG(VISIBILITY) AS AVG_VISIBILITY,
    AVG(DEW_POINT_TEMPERATURE) AS AVG_DEW_POINT_TEMPERATURE,
    AVG(SOLAR_RADIATION) AS AVG_SOLAR_RADIATION,
    AVG(RAINFALL) AS AVG_RAINFALL,
    AVG(SNOWFALL) AS AVG_SNOWFALL,
    AVG(RENTED_BIKE_COUNT) AS AVG_BIKE_COUNT
  FROM SEOUL_BIKE_SHARING
  GROUP BY SEASONS
  ORDER BY AVG_BIKE_COUNT DESC
  "
)
```


### Task 10: Use an implicit join across the WORLD_CITIES and the BIKE_SHARING_SYSTEMS tables to determine the total number of bikes avaialble in Seoul, plus the following city information about Seoul: CITY, COUNTRY, LAT, LON, POPULATION, in a single view.

Notice that in this case, the CITY column will work for the WORLD_CITIES table, but in general you would have to use the CITY_ASCII column.


```{r}
dbGetQuery(
  conn,
  "
  SELECT
    WC.CITY,
    WC.COUNTRY,
    WC.LAT,
    WC.LNG AS LON,
    WC.POPULATION,
    SUM(BS.BICYCLES) AS TOTAL_BIKES_AVAILABLE
  FROM
      WORLD_CITIES AS WC
  JOIN
      BIKE_SHARING_SYSTEMS AS BS
  ON
      WC.CITY = BS.CITY
  WHERE
      WC.CITY = 'Seoul'
  GROUP BY
      WC.CITY, WC.COUNTRY, WC.LAT, WC.LNG, WC.POPULATION
  "
)
```


### Task 11: Find all city names and coordinates with comparable bike scale to Seoul's bike sharing system

Find all cities with total bike counts between 15000 and 20000. Return the city and country names, plus the coordinates (LAT, LNG), population, and number of bicycles for each city.

Later we will ask you to visualize these similar cities on leaflet, with some weather data.

```{r}
dbGetQuery(
  conn,
  "
  SELECT
    WC.CITY,
    WC.COUNTRY,
    WC.LAT,
    WC.LNG,
    WC.POPULATION,
    BS.BICYCLES AS TOTAL_BIKES
  FROM
      WORLD_CITIES AS WC
  JOIN
      BIKE_SHARING_SYSTEMS AS BS
  ON
      WC.CITY = BS.CITY
  WHERE
      BS.BICYCLES BETWEEN 15000 AND 20000
  "
)
```

```{r}
dbDisconnect(conn)
```


## Analyse exploratoire des données avec Tidyverse et ggplot2

Dans cet atelier, vous utiliserez un notebook R pour effectuer une analyse exploratoire des données à l'aide de Tidyverse et des packages R ggplot2.

Vous commencerez par effectuer une préparation mineure des données sur l'ensemble de données SEOUL BIKE SHARING. Ensuite, vous générerez et explorerez quelques statistiques à partir de la dataframe résultante et ferez quelques observations. Enfin, vous générerez des graphiques informatifs à l'aide de la bibliothèque ggplot2.

Votre objectif principal est de recueillir des informations à partir de votre analyse exploratoire. Ces résultats feront partie de votre histoire et vous créerez votre présentation finale.

La visualisation est un outil très puissant pour mieux comprendre vos données. La visualisation peut également éclairer votre analyse en repérant les valeurs aberrantes et les comportements anormaux dans vos entités. Les boîtes à moustache (*boxplot*) sont informatifs à cet égard. Vous pouvez également obtenir des informations sur les tendances et anomalies claires qui peuvent être présentes dans une variable, simplement en la traçant directement. 

Commençons notre analyse exploratoire !

For reference, we include the Attribute Information for the `seoul_bike_sharing` dataset:

- DATE - format: "2017-12-01"
- RENTED_BIKE_COUNT - Count of bikes rented at each hour
- HOUR - Hour of the day
- TEMPERATURE - Celsius
- HUMIDITY - %
- WIND_SPEED - m/s
- VISIBILITY - 10m
- DEW_POINT_TEMPERATURE - Celsius
- SOLAR_RADIATION - MJ/m2
- RAINFALL - mm
- SNOWFALL - cm
- SEASONS -  "Autumn","Spring",..
- HOLIDAY - "Holiday", "No holiday"
- FUNCTIONING_DAY - "Yes", "No"


```{r}
# Importation des données
seoul_bike_sharing <- read_csv("seoul_bike_sharing_cleaned.csv")
head(seoul_bike_sharing)
```

```{r}
# Colonne 'DATE' au format date
seoul_bike_sharing$DATE <- as.Date(
  seoul_bike_sharing$DATE, 
  format = "%d/%m/%Y"
)
str(seoul_bike_sharing$DATE)
```

```{r}
# Colonne 'HOUR' au format Integer
seoul_bike_sharing$HOUR <- as.integer(seoul_bike_sharing$HOUR)
str(seoul_bike_sharing$HOUR)
```

```{r}
# Nombre de valeurs manquantes
sum(is.na(seoul_bike_sharing))
```

```{r}
# Résumé statistique
summary(seoul_bike_sharing)
```



```{r}
# Nombre d'observations qui tombent un jour férié
seoul_bike_sharing %>%
  filter(HOLIDAY == "Holiday") %>%
  nrow()
```

```{r}
# Calcule du total saisonnier des précipitations et des chutes de neige
seoul_bike_sharing %>%
  group_by(SEASONS) %>%
  summarise(
    total_rainfall = sum(RAINFALL),
    total_snowfall = sum(SNOWFALL)
  )
```

```{r}
# Nuage de points `RENTED_BIKE_COUNT` vs `DATE` coloriés par les saisons
ggplot(seoul_bike_sharing, aes(DATE, RENTED_BIKE_COUNT, color = SEASONS)) +
  geom_point(alpha=0.5) +
  labs(title = "Scatter Plot of Rented Bike Count Over Time",
       x = "Date",
       y = "Rented Bike Count") +
  theme_minimal()
```

En hiver, il y a moins de locations de vélos. C'est en été qu'il y a le plus grand nombre de vélos loués suivi respectivement par l'automne et le printemps. Donc on peut dire que la location de vélos est étroitement liée à la saison. Plus les conditions météorologiques sont agréables, plus les gens sont incités à utiliser le vélo.

```{r}
seoul_bike_sharing %>%
  group_by(SEASONS) %>%
  summarise(total_rented_bike = sum(RENTED_BIKE_COUNT))
```


```{r}
# Nuage de points `RENTED_BIKE_COUNT` vs `DATE` coloriés par l'heure
ggplot(seoul_bike_sharing, aes(DATE, RENTED_BIKE_COUNT, color = HOUR)) +
  geom_point(alpha=0.4) +
  scale_colour_viridis() +
  labs(title = "Scatter Plot of Rented Bike Count Over Time",
       x = "Date",
       y = "Rented Bike Count") +
  theme_minimal()
```

Let's visualize the correlation between RENTED_BIKE_COUNT and TEMPERATURE by SEASONS

```{r}
# Create the scatter plot with facets
ggplot(seoul_bike_sharing, aes(x = TEMPERATURE, y = RENTED_BIKE_COUNT, color = HOUR)) +
  geom_point(alpha = 0.2) +  
  scale_colour_viridis() +
  facet_wrap(~ SEASONS) +    # Create facets by SEASONS
  labs(title = "Correlation between Rented Bike Count and Temperature by Season",
       x = "Temperature",
       y = "Rented Bike Count") +
  theme_minimal()

```

```{r, fig.height=10}
# Create the boxplots with facets
ggplot(seoul_bike_sharing, aes(x = as.factor(HOUR), y = RENTED_BIKE_COUNT)) +
  geom_boxplot() +                    
  facet_wrap(~ SEASONS, ncol = 1) +   
  labs(title = "Boxplots of Rented Bike Count vs. Hour by Season",
       x = "Hour",
       y = "Rented Bike Count") +
  theme_minimal()
```



```{r}
daily_totals <- seoul_bike_sharing %>%
  group_by(DATE) %>%
  summarize(
    TOTAL_RAINFALL = sum(RAINFALL),
    TOTAL_SNOWFALL = sum(SNOWFALL)
  )

daily_totals
```

```{r}
# Create a line plot for daily total rainfall
ggplot(data = daily_totals, aes(x = DATE, y = TOTAL_RAINFALL)) +
  geom_line(color = "blue") +
  labs(title = "Daily Total Rainfall",
       x = "Date",
       y = "Rainfall (mm)") +
  theme_minimal()

# Create a line plot for daily total snowfall
ggplot(data = daily_totals, aes(x = DATE, y = TOTAL_SNOWFALL)) +
  geom_line(color = "red") +
  labs(title = "Daily Total Snowfall",
       x = "Date",
       y = "Snowfall (cm)") +
  theme_minimal()
```

Let's determine determine how many days had snowfall:

```{r}
# Filter for days with nonzero snowfall
snowfall_days <- daily_totals[daily_totals$TOTAL_SNOWFALL > 0, ]

# Count the unique dates with snowfall
num_snowfall_days <- length(unique(snowfall_days$DATE))

# Print the result
cat("Number of days with snowfall:", num_snowfall_days, "\n")
```

# Partie 4 : Analyse prédictive. Prédire la demande de vélos en libre-service à l'aide de modèles de régression

Maintenant que vous avez effectué une analyse exploratoire et visuelle des ensembles de données sur la demande de vélos en libre-service et obtenu quelques informations préliminaires sur les attributs, il est temps de créer, d'évaluer et d'affiner plusieurs modèles prédictifs et de trouver le modèle le plus performant pour prédire le nombre horaire de location de vélos.

## Création d'un modèle de régression linéaire de référence

### Importation des données

```{r}
# importation des données
seoul_bike_sharing <- read_csv("seoul_bike_sharing_cleaned.csv")
head(seoul_bike_sharing)
```

We won't be using the `DATE` column, because 'as is', it basically acts like an data entry index. (However, given more time, we could use the `DATE` colum to create a 'day of week' or 'isWeekend' column, which we might expect has an affect on preferred bike rental times.) We also do not need the `FUNCTIONAL DAY` column because it only has one distinct value remaining (`YES`) after missing value processing.

```{r}
seoul_bike_sharing <- seoul_bike_sharing %>%
  select(-c(DATE, FUNCTIONING_DAY))
```


### Ensemble d'entraînement et de test

```{r}
set.seed(21)
# Objet Split
bike_split <- initial_split(
  data = seoul_bike_sharing,
  prop = 3/4
)

# Données d'entrainement (75%)
train_data <- training(bike_split)
dim(train_data)

# Données d'évaluation du modèle final (25%)
test_data <- testing(bike_split)
dim(test_data)
```
### Construction d'un modèle de régression linéaire en utilisant uniquement des variables météorologiques

Comme vous pouvez l'imaginer, les conditions météorologiques peuvent affecter les décisions des gens en matière de location de vélos. Par exemple, par une journée froide et pluvieuse, vous pouvez choisir un autre moyen de transport comme un bus ou un taxi. Par une belle journée ensoleillée, vous souhaiterez peut-être louer un vélo pour un trajet de courte distance.

Construisons donc un modèle de régression linéaire en utilisant les variables suivantes :

- `TEMPERATURE` - Temperature in Celsius
- `HUMIDITY` - Unit is `%`
- `WIND_SPEED` - Unit is `m/s`
- `VISIBILITY` - Multiplied by 10m
- `DEW_POINT_TEMPERATURE` - The temperature to which the air would have to cool down in order to reach saturation, unit is Celsius
- `SOLAR_RADIATION` - MJ/m2
- `RAINFALL` - mm
- `SNOWFALL` - cm

```{r}
lm_model_weather <- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

lm_model_weather
```
```{r}
formula <- RENTED_BIKE_COUNT ~ TEMPERATURE + HUMIDITY + WIND_SPEED + VISIBILITY +
            DEW_POINT_TEMPERATURE + SOLAR_RADIATION + RAINFALL + SNOWFALL
# Fit the model with the formula and training data
lm_model_weather_fit <- fit(lm_model_weather, formula, data = train_data)

print(lm_model_weather_fit$fit)
```
```{r}
broom::tidy(lm_model_weather_fit$fit)
```
### Construction d'un modèle de régression en utilisant toutes les variables

```{r}
lm_model_all <- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

lm_model_all_fit <- fit(
  lm_model_all, 
  formula = RENTED_BIKE_COUNT ~ ., 
  data = train_data
)

print(lm_model_all_fit$fit)

broom::tidy(lm_model_all_fit$fit)
```

### Évaluation du modèle et identification des variables importantes

```{r}
# Make predictions using lm_model_weather_fit
test_results_weather <- predict(lm_model_weather_fit, test_data) %>%
  mutate(
    truth = test_data$RENTED_BIKE_COUNT,
    estimate = .pred
  ) %>%
  select(-.pred)
head(test_results_weather, 10)

# Make predictions using lm_model_all_fit
test_results_all <- predict(lm_model_all_fit, new_data = test_data) %>%
  mutate(
    truth = test_data$RENTED_BIKE_COUNT,
    estimate = .pred
  ) %>%
  select(-.pred)
head(test_results_all, 10)
```


```{r}
# Calculate R-squared for lm_model_weather_fit
rsq_weather <- rsq(
  data = test_results_weather, 
  truth = truth,
  estimate = estimate
) %>%
  pull(.estimate)

# Calculate RMSE for lm_model_weather_fit
rmse_weather <- rmse(
  data = test_results_weather, 
  truth = truth,
  estimate = estimate
) %>%
  pull(.estimate)

# Calculate R-squared for lm_model_all_fit
rsq_all <- rsq(
  data = test_results_all, 
  truth = truth,
  estimate = estimate
) %>%
  pull(.estimate)

# Calculate RMSE for lm_model_all_fit
rmse_all <- rmse(
  data = test_results_all, 
  truth = truth,
  estimate = estimate
) %>%
  pull(.estimate)

# Print the results
cat("R-squared (lm_model_weather_fit):", rsq_weather, "\n")
cat("RMSE (lm_model_weather_fit):", rmse_weather, "\n")

cat("R-squared (lm_model_all_fit):", rsq_all, "\n")
cat("RMSE (lm_model_all_fit):", rmse_all, "\n")

```

D'après ces tableaux, nous constatons que le modèle `lm_model_all_fit` est plus performant que `lm_model_weather_fit`. Cela signifie que l'utilisation de toutes les variables (variables météorologiques et date-heure) dans le modèle génère de meilleurs résultats de prévision.

```{r}
# Get the coefficients from lm_model_all_fit
coefficients <- lm_model_all_fit$fit$coefficients

# Convert coefficients to a data frame
coefficients_df <- data.frame(
  Variable = names(coefficients),
  Coefficient = coefficients
)

# Filter out the 'Intercept' row
coefficients_df <- coefficients_df %>%
  filter(Variable != "(Intercept)")

# Sort the coefficients in descending order
coefficients_df <- coefficients_df %>%
  arrange(desc(Coefficient))

# Create a bar plot of coefficients
ggplot(coefficients_df, aes(x = reorder(Variable, -Coefficient), y = Coefficient)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "blue") +
  labs(title = "Coefficients of lm_model_all_fit",
       x = "Variable",
       y = "Coefficient") +
  coord_flip() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Affiner les modèles de régression de base

```{r}
# Données
bike_sharing_df <- read_csv("seoul_bike_sharing_cleaned.csv")
bike_sharing_df <- bike_sharing_df %>%
  select(-DATE, -FUNCTIONING_DAY)

head(bike_sharing_df)
```


```{r}
# Données d'entrainement et de test
set.seed(1234)
data_split <- initial_split(bike_sharing_df, prop = 4/5)
train_data <- training(data_split)
test_data <- testing(data_split)
```


```{r}
# Définition d'un modèle de régression
lm_spec <- linear_reg() %>%
  set_engine("lm") %>% 
  set_mode("regression")

lm_spec
```
### Ajout des termes polynomiaux

Les modèles de régression linéaire sont les modèles les plus appropriés pour capturer les corrélations linéaires entre les variables. Cependant, dans les données réelles, de nombreuses relations peuvent être non linéaires.

Par exemple, la correlation entre `RENTED_BIKE_COUNT` et `TEMPERATURE` ne semble pas être linéaire :

```{r}
ggplot(data = train_data, aes(RENTED_BIKE_COUNT, TEMPERATURE)) + 
    geom_point() 
```

```{r}
# Plot the higher order polynomial fits
ggplot(data=train_data, aes(RENTED_BIKE_COUNT, TEMPERATURE)) + 
    geom_point() + 
    geom_smooth(method = "lm", formula = y ~ x, color="red") + 
    geom_smooth(method = "lm", formula = y ~ poly(x, 2), color="yellow") + 
    geom_smooth(method = "lm", formula = y ~ poly(x, 4), color="green") + 
    geom_smooth(method = "lm", formula = y ~ poly(x, 6), color="blue")
```

```{r, eval=FALSE, echo=TRUE}
# Get the coefficients and variable names from lm_model_all_fit
coefficients <- lm_model_all_fit$fit$coefficients
variable_names <- names(coefficients)

# Exclude the "(Intercept)" coefficient from variable_names
variable_names <- variable_names[variable_names != "(Intercept)"]

# Identify the important variables with larger coefficients
# For example, let's consider variables with coefficients greater than a threshold
threshold <- 0.1  # Adjust this threshold as needed
important_variables <- variable_names[abs(coefficients[-1]) > threshold]  # Exclude "(Intercept)"

# Select the important variables from the training data
selected_variables <- train_data[, important_variables]

# Create polynomial terms for the selected variables (e.g., quadratic terms)
selected_variables_poly <- poly(selected_variables, degree = 2)

# Combine the polynomial terms with the original dataset
train_data_poly <- cbind(train_data, selected_variables_poly)

# Fit the linear regression model with polynomial terms (excluding "(Intercept)")
lm_poly <- lm(RENTED_BIKE_COUNT ~ . - 1, data = train_data_poly)
```

# Partie 5 : Création d'une application de prévision de la demande de vélos en libre-service avec R Shiny et Leaflet

The code of `ui.R`:

```{r, eval=FALSE, echo=TRUE}
# Load required libraries
require(leaflet)

# Create a RShiny UI
shinyUI(
  fluidPage(padding=5,
            titlePanel("Bike-sharing demand prediction app"), 
            # Create a side-bar layout
            sidebarLayout(
              # Create a main panel to show cities on a leaflet map
              mainPanel(
                # leaflet output with id = 'city_bike_map', height = 1000
                leafletOutput("city_bike_map", height = 1000)
              ),
              # Create a side bar to show detailed plots for a city
              sidebarPanel(
                # select drop-down list to select city
                selectInput("selected_city", "Select a City", choices = cities_df$CITY_ASCII),
                # Add more UI elements as needed
                plotOutput("temperature_trend_plot", height = 400)
              )
            )
  )
)

```


The code of `server.R`:

```{r, echo=TRUE, eval=FALSE}
# Load required libraries
require(shiny)
require(leaflet)
require(tidyverse)
require(plotly)

# Import the functions from model_prediction.R
source("model_prediction.R")

test_weather_data_generation<-function(){
  #Test generate_city_weather_bike_data() function
  city_weather_bike_df<-generate_city_weather_bike_data()
  stopifnot(length(city_weather_bike_df)>0)
  print(head(city_weather_bike_df))
  return(city_weather_bike_df)
}

city_weather_bike_df <- test_weather_data_generation()

# Load the cities data from selected_cities.csv
cities_df <- read.csv("selected_cities.csv")

# Load the trained regression model coefficients
model_coefficients <- read.csv("model.csv")

# Create a RShiny server
shinyServer(function(input, output){
  
  # Define color factor
  color_levels <- colorFactor(c("green", "yellow", "red"), 
                              levels = c("small", "medium", "large"))
  
  # Create a reactive function to filter data based on selected city
  filtered_data <- reactive({
    if (is.null(input$selected_city)) {
      return(NULL)
    }
    city_weather_bike_df %>%
      filter(CITY_ASCII == input$selected_city)
  })
  
  # Create a Leaflet map
  output$city_bike_map <- renderLeaflet({
    leaflet(data = filtered_data()) %>%
      addTiles() %>%
      addCircleMarkers(
        radius = ~sqrt(BIKE_PREDICTION) * 5,
        color = ~color_levels(BIKE_PREDICTION_LEVEL),
        fillOpacity = 0.7,
        popup = ~DETAILED_LABEL
      )
  })
  
  # Load the model coefficients and create a reactive function for predictions
  model_coeff <- reactive({
    if (is.null(input$selected_city)) {
      return(NULL)
    }
    model_coefficients %>%
      filter(Variable != "Intercept") %>%
      arrange(desc(Coef)) %>%
      select(Variable) %>%
      pull()
  })
  
  # Create a reactive function for making predictions based on selected city
  bike_predictions <- reactive({
    if (is.null(input$selected_city)) {
      return(NULL)
    }
    city_weather_bike_df %>%
      filter(CITY_ASCII == input$selected_city) %>%
      select({{ model_coeff() }}) %>%
      summarise_all(sum) %>%
      transmute(BIKE_PREDICTION = as.numeric(.)) %>%
      mutate(BIKE_PREDICTION_LEVEL = calculate_bike_prediction_level(BIKE_PREDICTION))
  })
  
  selected_city_data <- reactive({
    filtered_data() %>%
      mutate(HOUR = as.POSIXlt(FORECASTDATETIME)$hour)
  })
  
  output$temperature_trend_plot <- renderPlot({
    
    if (!is.null(selected_city_data())) {
      # Plot temperature trend for the selected city
      ggplot(selected_city_data(), aes(x = HOUR, y = TEMPERATURE)) +
        geom_line() +
        geom_point() +
        geom_text(aes(label = TEMPERATURE), hjust = 1.2, vjust = 0.5) +
        labs(title = "Temperature Trend for the Next 5 Days",
             #x = "Date and Time",
             x = "Hour",
             y = "Temperature (°C)") +
        theme_minimal()
    }
  })
  
})

```









